Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-11-28 inst/paper/main.tex
cites to Moller

Most dynamic languages provide their users with a facility to
transform unstructured text into executable code, and to evaluate that
code. In this paper we refer to this reflective facility of a language
as \eval bowing to its origins in LISP in 1956~\cite{lisp}. \Eval has
been much maligned over the years. In computing lore, it is as close
to a boogeyman as it gets. Yet, for McCarthy, \eval was simply the way
to write down the definition of LISP, he was surprised that someone
coded it up and offered it to end users.  Since then, reflective
facilities have been used to parameterize programs over code patterns
that can be provided after the program is written.  The presence of
such a feature in a language is a hallmark of dynamism as it is a form
of delayed binding as what any particular call to \eval will do is
only know at program run-time when that particular call site is
evaluated.

\vspace{2mm}\noindent\emph{Trouble in Paradise.} Reflective facilities
hinder most attempt to reason about the code using them, as well as to
apply meaning-preserving program transformations. In practice this
means that static analysis techniques will loose so much precision as
to become pointless and that anything but the most trivial, local,
compiler optimizations will be unsound. Furthermore the addition of
new code --- code that could have been obtained from a network
connection --- as the program is running is a security vulnerability
waiting to happen.

To further illustrate the challenges at hand, consider the interaction
between a static analysis tool and a dynamic language. An abstract
interpretation-based program analyzer computes an over-approximation
of the set of possible behaviors exhibited by the system under
study~\cite{cc77}. The presence of a reflective operation entails that
the operation may have \emph{any} behavior that can be expressed in
the target language; i.e. as if we replaced it by any legal sequence
of instructions. As dynamic languages tend to be permissive, the
analysis has to, for example, assume that many (and sometimes all)
functions in scope may have been redefined, e.g. that \texttt{`+`} now
opens an Internet connection or something equally surprising. This
means that a single occurrence of \eval causes the static analyzer to
loose all information about the program state and the meaning of
identifiers. This loss of precision can sometimes be mitigated by
analyzing the string passed to \eval~\cite{} to bound its possible
behavior but when the string comes from outside the program not much
can be done. A frustrated group researchers argued giving up on
soundness and, instead, under-approximating dynamic features
(soundiness)~\cite{soundy}. In their words ``a practical analysis,
therefore, may pretend that \eval does nothing, unless it can
precisely resolve its string argument at compile time.'' Alas,
assuming that \eval does not have side-effects, or that side-effects
will not affect the results of the analysis, may be unduly optimistic.

@@

\vspace{2mm}\noindent\emph{The here and now.}  In this study, we
investigate the usage of \eval in programs written in the R
programming language. R is language designed by statisticians for
applications in data science~\cite{r,R96}. What makes looking at R after
JavaScript interesting is that, while both languages are dynamic, they
are quite different. While one can program in an object-oriented style
in R, like in JavaScript, R is primarily a lazy, untyped, functional
language.  JavaScript was designed to run untrusted code in browser,
while R is used for statistical computing on desktops. JavaScript is a
general purpose language used by a wide community of programmers;
while R is used for scientific computing by data scientists and domain
experts with, often, limited programming experience. One can
distinguish between library implementers, developers with some
programming experience and a working knowledge of R, and end-users,
who are typically not expert programmers and often have only a cursory
knowledge of the language.\footnote{Consider that R evaluates function
argument lazily, just like Haskell. We informally surveyed end-users,
including computer scientists, and did not find a single user aware of
this fact. Library developers, on the other hand, know about laziness
and program defensively around it.} Our goal is thus to highlight the
differences in usage between JavaScript and R, and try to explain
those differences in terms of language features, application domain
and programmer experience.

\vspace{2mm}\noindent\emph{The What and How.}  One significant benefit
of choosing R is that every package in the CRAN repository is curated
and comes with examples of typical usage. This gives us a large code
base that we can analyze dynamically.  To observe \eval we built a
two-level monitoring infrastructure: we can monitor R programs by
instrumentation --- this gives us access to many user-visible properties
of R programs --- but we can also monitor the inner working of the R
interpreter --- this allows us to capture details of execution that are
not exposed at the source level. Dynamic analysis is of course limited
in that it can only observe behaviors that are triggered by the
particular inputs passed to the program. Luckily, CRAN libraries come
with multiple input datasets and use-cases.

@@

Our infrastructure is publicly available, released in open source, and
our code, and data will be submitted for artifact evaluation.

@@

The expressive power of \eval depends on a number of choices made by
the programming language designers. The two axes that are the most
relevant here are:
\begin{itemize}
\item {\bf Scoping:} What environment does the string passed to \eval
  execute in? JavaScript and R evaluate the expression in the current
  scope. This means that local variables and parameters are exposed to
  \eval, breaking the abstraction boundary of the enclosing function.
  Julia takes a more restrictive approach by running \eval at the
  ``top level'', i.e. in the global environment, outside of any
  function's environment. JavaScript has a global \eval that behaves
  like Julia, it also has a ``strict mode'' in which \eval is allowed
  to access local variables, but fobidden from injecting new variables
  in the local environment. In Java, one can implement \eval but the
  expression will execute in new environment.\footnote{While \eval
  used to be available only in interpreted languages, the advent of
  just-in-time compilation gradually lifted this restriction. To
  implement \eval in Java requires wrapping its input in a static
  method of an anonymous class, generating bytecode for that class,
  invoking the class loader to install the code, and then using
  reflection to call the new method. }
\item {\bf Reflective API:} What other reflective operations are
  exposed to user code? JavaScript does provides few reflective
  functions (other than being able to enumerate the properties of an
  object and to access properties with strings), the string passed to
  \eval is thus limited to what can be expressed in JavaScript. R, on
  the other hand, has a rich reflective interface. For example, R
  allows user code to walk the call stack and arbitrarily add and
  delete local variables in any environment.
\end{itemize}

\noindent
These design choices impact our ability to reason about code. More
expressive power for \eval translates directly into additional
restriction for program analysis and transformation tools. The reason
Julia has restricts \eval to execute in top level environment is that
they wanted to preserve their compiler's ability to optimize
functions~\cite{oopsla18a}. What they wanted to prevent was the case
where \eval changed the type of local variables as this would require
recompilation of the method (or prevent optimization). Julia even adds
a versoning mechanism, called world age, to ensure that code added
during an \eval does not invalidate optimizations such as
inlining~\cite{oopsla20a}. More permissive languages such as R are
much harder to deal with from a compiler's perspective.

\medskip\noindent{\it Eval in R.} There are four reflective functions
that take abstract syntax trees (called expressions), parse it as
code, and evaluate it. These functions are \c{eval}, \c{evalq},
\c{eval.parent}, and \c{local}. Unlike other languages, R's \eval does
not take a string, but rather an expression. As this is slightly more
involved, it may discourage some casual users. There are two ways to
obtain an expression. Given a string, the \c{parse()} function retuns
an unevaluated expression. The other way is to use the function
\c{substitute(e,envir)}, it yields the parse tree of the expression
\c{e} after performing substitutions defined by the bindings in the
\c{envir} parameter.

@@

Richards et al.~\cite{ecoop11} provided the first large-scale study of
the runtime behavior of \eval in JavaScript. They dynamically analyzed
a corpus of the 10,000 most popular websites with an instrumented web
browser to gather execution traces.  They show that \eval is pervasive
with 82\% of the most popular websites using it. The reasons for its
use include the desire to load code on demand, deserialization of JSON
dataq and lightweight meta-programming to customize web pages.  While
many uses were legitimate, just as many were unnecessary and could be
replaced with equivalent and safer code.  They categorized inputs to
\eval so as to cover the vast majority of input strings.  Restricting
themselves to \eval in which all named variables refer to the global
scope, many patterns could be replaced by more disciplined
code~\cite{oopsla12b}.  The work did not measure code coverage, so the
numbers presented are a lower bound on the possible behaviors.
Furthermore, JavaScript usage in 2011 is likely different from today,
e.g. Node.js was not covered by Richards.  More details about dynamic
analysis of JavaScript can be found in~\cite{liang}.

>>>>

Most dynamic languages provide their users with a facility to
transform unstructured text into executable code and evaluate that
code. We refer to this reflective facility as \eval bowing to its
origins in LISP, all the way back in 1956~\cite{lisp}. \Eval has been
much maligned over the years. In computing lore, it is as close to a
boogeyman as it gets. Yet, for McCarthy, \eval was simply the way to
write down the definition of LISP, he was surprised that someone coded
it up and offered it to end users. Since then, reflective facilities
have been used to parameterize programs over code patterns that can be
provided after the program is written. The presence of such a feature
in a language is a hallmark of dynamism; it is a form of delayed
binding as the behavior of any particular call to \eval will only be
known when the program is run and that particular call site is
evaluated.

\vspace{2mm}\noindent\emph{Trouble in Paradise.} Reflective facilities
hinder most attempts to reason about, or apply meaning-preserving
transformation to, the code using them. In practice, \eval causes
static analysis techniques to loose so much precision as to become
pointless. For compilers, anything but the most trivial, local,
optimizations are unsound after a use of \eval. Furthermore, the
addition of arbitrary code --- code that could have been obtained from
a network connection --- as a program is running is a security
vulnerability waiting to happen. To illustrate these challenges,
consider the interaction of a static analysis tool with a dynamic
language. An abstract interpretation-based program analyzer computes
an over-approximation of the set of possible behaviors exhibited by
the program under study~\cite{cc77}. A reflective may have \emph{any}
behavior that can be expressed in the target language; i.e. \eval can
be replaced by any legal sequence of instructions. As dynamic
languages tend to be permissive, the analysis has to, for example,
assume that many (or all) functions in scope may have been redefined,
e.g. that \texttt{`+`} now opens a network connection or something
equally surprising. A single occurrence of \eval causes the static
analyzer to loose all information about program state and
meaning of identifiers. This loss of precision can sometimes be
mitigated by analyzing the string argument~\cite{moller03} to bound its
possible behavior but when the string comes from outside the program
not much can be done. A frustrated group researchers argued giving up
on soundness and, instead, under-approximating dynamic features
(soundiness)~\cite{soundy}. In their words ``a practical analysis,
therefore, may pretend that \eval does nothing, unless it can
precisely resolve its string argument at compile time.'' Alas,
assuming that \eval does not have side-effects, or that side-effects
will not affect the results of the analysis, may be unduly optimistic.

@@

\vspace{2mm}\noindent\emph{The Here and Now.} In this study, we
investigate the usage of \eval in programs written in the R
programming language. R is language designed by statisticians for
applications in data science~\cite{r,R96}. What makes looking at R
after JavaScript interesting is that, while both languages are
dynamic, they are quite different. While one can program in an
object-oriented style in R, like in JavaScript, R is primarily a lazy,
untyped, functional language. JavaScript was designed to run untrusted
code in browser, while R is used for statistical computing on
desktops. JavaScript is a general purpose language used by a wide
community of programmers; while R is used for scientific computing by
data scientists and domain experts with, often, limited programming
experience. One can distinguish between library implementers,
developers with some programming experience and a working knowledge of
R, and end-users, who are typically not expert programmers and often
have only a cursory knowledge of the language.\footnote{Consider that
R evaluates function argument lazily, just like Haskell. We informally
surveyed end-users, including computer scientists, and did not find a
single user aware of this fact. Library developers, on the other hand,
know about laziness and program defensively around it.} Our goal is
thus to highlight the differences in usage between JavaScript and R,
and try to explain those differences in terms of language features,
application domain and programmer experience. Hopefully some of our
observations will generalize to other languages.

\vspace{2mm}\noindent\emph{The What and How.} One significant benefit
of choosing R is that every package in the CRAN repository is curated
and comes with examples of typical usage. This gives us a large code
base that we can analyze dynamically. To observe \eval we built a
two-level monitoring infrastructure:\footnote{Our infrastructure is
open source and publicly available, our anonymized code is at {\tt
  http://url.com}, a more complete artifact will be submitted to the
artifact evaluation committee. } we can monitor R programs by
instrumentation --- this gives us access to many user-visible
properties of R programs --- but we can also monitor the inner working
of the R interpreter --- this allows us to capture details of
execution that are not exposed at the source level. Dynamic analysis
is of course limited in that it can only observe behaviors that are
triggered by the particular inputs passed to the program. Luckily,
CRAN libraries come with multiple input datasets and use-cases.

@@

\vspace{2mm}\noindent\emph{Why do we Eval?} {\bf A short summary of the
results should go here.}

@@

The expressive power of \eval depends on choices made by the language
designers along the following two axes:
\begin{itemize}
\item {\bf Scoping:} What environment does the string passed to \eval
  execute in? JavaScript and R evaluate it in the current environment,
  thus exposing local variables and parameters to \eval and breaking
  the abstraction boundary of the enclosing function. Julia is more
  restrictive, it runs \eval at the ``top level'', in the global
  environment. JavaScript has a global \eval that behaves like Julia,
  it also has a ``strict mode'' in which \eval is allowed to access
  local variables, but fobidden from injecting new variables in the
  local environment. In Java, one can implement \eval but the
  expression will execute in new environment.\footnote{While \eval
  used the purview of interpreted languages, the advent of
  just-in-time compilation lifted this restriction. To implement \eval
  in Java requires wrapping its input in a static method of a new
  anonymous class, generating bytecode for that class, invoking the
  class loader to install that code, and finally reflectively calling
  the new method. }
\item {\bf Reflective API:} What other reflective operations are
  exposed to user code? JavaScript provides few reflective functions
  (other than enumeration of the properties of an object and access to
  properties with strings), \eval is thus limited to what can be
  expressed in JavaScript code. R, on the other hand, has a rich
  reflective interface. For example, R allows user code to walk the
  call stack and arbitrarily add and delete local variables in any
  environment. This reflective API means that there are few limits to
  what can do.
\end{itemize}

\noindent
These design choices impact our ability to reason about code when
reflection is used. A more expressive \eval translates directly into
additional restrictions for program analysis and transformation tools.
The reason Julia restricts \eval to the top level environment is to
preserve the compiler's ability to optimize
functions~\cite{oopsla18a}. What the Julia designers wanted was to
prevent \eval from modifying the type of local variables as this would
require recompilation of the method. Julia even adds a versoning
mechanism, called world age, to ensure that code added during an \eval
does not invalidate optimizations such as inlining~\cite{oopsla20a}.
More permissive languages such as R are much harder to deal with from
a compiler's perspective.

\medskip\noindent{\it Eval in R.} There are four reflective functions
that take abstract syntax trees (called expressions), parse them as
code, and evaluate the result. These functions are \c{eval},
\c{evalq}, \c{eval.parent}, and \c{local}. Taking an expression rather
than a string, may discourage some casual users as manipulating
expressions is slightly more involved.\footnote{The rationale for
\eval taking expressions is explained by the main use case for \eval,
namely extending the language by reinterpretation of promises; R gives
access to the code of a promise as an expression.} There are two ways
to obtain an expression. Given a string, the \c{parse()} function
retuns an unevaluated expression. The other way is to use the function
\c{substitute(e,envir)}, it yields the parse tree of the expression
\c{e} after performing substitutions defined by the bindings in the
\c{envir} parameter.

@@

Richards et al.~\cite{ecoop11} provided the first large-scale study of
the runtime behavior of \eval in JavaScript. They dynamically analyzed
a corpus of the 10,000 most popular websites with an instrumented web
browser to gather execution traces. They show that \eval is pervasive
with 82\% of the most popular websites using it. The reasons for its
use include the desire to load code on demand, deserialization of JSON
dataq and lightweight meta-programming to customize web pages. While
many uses were legitimate, just as many were unnecessary and could be
replaced with equivalent and safer code. They categorized inputs to
\eval so as to cover the vast majority of input strings. Restricting
themselves to \eval in which all named variables refer to the global
scope, many patterns could be replaced by more disciplined
code~\cite{oopsla12b, moller12}. The work did not measure code
coverage, so the numbers presented are a lower bound on the possible
behaviors. Furthermore, JavaScript usage in 2011 is likely different
from today, e.g. Node.js was not covered by Richards. More details
about dynamic analysis of JavaScript can be found in~\cite{liang}.
